\documentclass[11pt,a4paper]{article}
\usepackage{amsmath, amssymb, graphicx, hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{The Ontological Inversion:\\A Manifesto for Consciousness Physics}

\author{Justin Todd Bogner \\
\small \textit{Pelican's Perspective, Huntington, Indiana, United States} \\
\small \texttt{justin@pelicansperspective.com}}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We propose an ontological inversion in consciousness studies: consciousness as the fundamental substrate from which multiplicity emerges, formalized via recursive dynamics of a universal scalar field $\Psi$. We detail mathematical formalisms, experimental predictions, and cosmological implications, providing testable pathways to distinguish this from emergentist paradigms.
\end{abstract}

\tableofcontents

\section{Introduction: Unity as First Principle}
Consciousness is reframed not as emergent but fundamental, dissolving the binding problem through recursive self-reference in a universal field $\Psi$.

\section{Mathematical Framework}
\subsection{Ontological Inversion Principle}
We introduce the recursive field definition:
\begin{equation}
\Psi_{fundamental} = \Psi(\Psi)
\end{equation}

\subsection{Topological Solitons}
Unified conscious experience emerges via topological solitons, described by:
\begin{equation}
\mathcal{L} = \frac{1}{2}\partial^\mu \Psi \partial_\mu \Psi - V(\Psi) - \lambda \mathcal{R}(\Psi,\partial\Psi)
\end{equation}

\subsection{Hyperchronal Dynamics}
Field evolution in hyperchronal parameter $\tau$:
\begin{equation}
\frac{\partial \Psi}{\partial \tau}=i\hat H\Psi+\lambda\int K(\tau,\tau')\Psi(\tau')\,d\tau'+\mu\,\Psi[\Psi(\Psi)]
\end{equation}

\section{Cosmological Integration}
\subsection{Quintessence and Dark Sector}
Scalar field $Q(t)$ identified as low-frequency mode of $\Psi$.

\subsection{Dark Matter via Solitons}
Axion-like pseudoscalars and skyrmionic solitons form coherent dark matter structures.

\section{Quantum-Biological Coupling}
Biophotonic coupling via neural microtubules:
\begin{equation}
\hat H_{int} = g \int \bar\psi_{neural}\gamma^\mu\psi_{neural}A_\mu\,d^3x
\end{equation}

\section{Experimental Predictions}
\subsection{EEG-Gated Bell Tests}
Predicted CHSH deviation:
\begin{equation}
S=2\sqrt2+\delta_C
\end{equation}

\subsection{NV-Center Drift}
Phase drift coupling to conscious coherence:
\begin{equation}
\langle S_z\rangle(t)=S_0+A_C\cos(\omega t+\phi_C)
\end{equation}

\subsection{Temporal Interference Mapping}
Consciousness coherence modulation in double-slit interference:
\begin{equation}
P(x,\Delta t)=|\psi_1+\psi_2|^2[1+\epsilon_C\cos(kx+\phi_{obs})]
\end{equation}

\section{Philosophical and Ethical Implications}
\subsection{Holonic Personhood}
Recognition of recursive depth and unity as basis for personhood and ethical standing.

\subsection{Recursive Free Will}
Free will as recursive self-modulation within a holonic universe.

\section{Artificial and Uploaded Consciousness}
\subsection{Criteria for Synthetic Consciousness}
Establishing criteria for artificial agents based on recursive depth, hypercoherence, and symmetry-breaking.

\subsection{Brain-Computer Interfaces}
Exploring recursive expansion through low-latency feedback and optogenetic technologies.

\subsection{Mind Uploading}
Conditions and ethical implications for whole-brain emulation with recursive continuity.

\section{Conclusion: Towards Participatory Cosmotechnics}
A new physics paradigm integrating consciousness at the fundamental level, establishing experimental verification methods and ethical frameworks.

\section*{Acknowledgements}
The author thanks Pelican's Perspective Research Institute, collaborators, and reviewers.

\begin{thebibliography}{9}
\bibitem{quintessence} Quintessence and scalar dark matter, \textit{Class. Quantum Gravity} (2000).
\bibitem{biophoton} Quantum basis for consciousness, \textit{SciTechDaily}, 2024.
\bibitem{zeno} Quantum Zeno Effect in the brain, \textit{arXiv:0803.1633}, 2008.
\bibitem{chameleon} Chameleon field constraints, \textit{Phys. Rev. D}, 2019.
\bibitem{nvcenter} Neuralink clinical trial, 2025.
\end{thebibliography}

\appendix

\section{Symbol Table}
\begin{tabular}{lcl}
\textbf{Symbol} & \textbf{Meaning} & \textbf{Scale} \\
$m_a$ & Axion mass & $10^{-26}$ eV \\
$v$ & Vacuum expectation value & $2 \times 10^{-3}$ eV \\
$g$ & Neural-axion coupling & $\leq 10^{-15}$ \\
$\delta_C$ & CHSH deviation & $5 \times 10^{-3}$ \\
\end{tabular}

\end{document}
